Preregistration analysis, Cost 2 study

### 2019-05-08 Cost2/Cost2_prereg_remake.Rmd
matching styles between Cost 2 and 3 reports

---
title: "Effort and Delay Discounting in a Foraging Environment"
author: "Claudio Toro Serey & Joseph T. McGuire"
output:
  html_document: default
  pdf_document: default
  fig_caption: yes
  word_document: default
---

# Pre-registration follow-up

@@

## Research Questions

#### 3.1.	Do decision makers in foraging environments integrate information about delay durations and reward amounts to produce reward-maximizing behavior?

#### 3.2.	Main question: Are decisions affected differently by equivalent time periods of pure delay, cognitive effort, physical effort, and non-effortful task engagement? Do these four conditions involve different levels of subjective costs?

#### 3.3.	How can we best computationally model the perceived cost of time and effort to predict choices in each condition?

## Hypotheses

#### 4.1.	Single-option, accept/reject decisions will be influenced by within-subject manipulations of reward magnitude and associated delay duration in line with a theoretical reward-maximizing strategy.
 **4.1.1.**	Participants will more frequently accept high-reward prospects than low-reward prospects.

@@

#### 4.2.	Prospect acceptance rates will differ across four between-subject conditions, in which the delays associated with rewards (a) are unfilled, (b) include a cognitive effort requirement, (c) include a physical effort requirement, or (d) require a trivial level of physical effort but have matched visual stimuli to the physical effort condition.
  **4.2.1.**	In the cognitive effort condition, overall acceptance rates will be greater than in the unfilled-delay condition.

@@

#### 4.3.	Choices will be well fit by a computational model in which the subjective opportunity cost of time is free to vary across the four between-subject conditions.
  **4.3.1.**	Participants will display stable preferences, meaning that the reward amounts they accept in a given timing condition will be similar throughout the experiment.

@@

## Analyses

@@

``` {r Basic Prepping, echo=FALSE}  
# # params for reference
# handling <- c(2, 10, 14)
# cond <- c("Wait", "Cognitive", "Physical", "Easy")
# # create data frame with summary stats
# tempW <- summarizeData()
# tempC <- summarizeData(Data = cogData, nSubs = nSubsC, type = "Cognitive")
# tempP <- summarizeData(Data = physData, nSubs = nSubsP, type = "Physical")
# tempE <- summarizeData(Data = easyData, nSubs = nSubsE, type = "Easy")
# summaryData <- rbind(tempW,
#                      tempC,
#                      tempP,
#                      tempE)
# colnames(summaryData) <- c("Group", "propComplete", "Earnings", "RT", "forcedTravel")
# summaryData$Cols <- c(rep(cols[1],nSubsW), rep(cols[2],nSubsC), rep(cols[3],nSubsP), rep(cols[4],nSubsE))
# rm(tempW, tempC, tempP, tempE)
# # Indexes for groups
# indxWait <- seq(nSubsW)
# indxWaitLong <- seq(nSubsW * 9) # 9 being the number of combinations
# indxCog <- (nSubsW + 1):(nSubsW + nSubsC)
# indxCogLong <- ((nSubsW * 9) + 1):((nSubsW * 9) + (nSubsC * 9))
# indxPhys <- ((nSubsW + nSubsC) + 1):(nSubsW + nSubsC + nSubsP)
# indxPhysLong <- ((nSubsW * 9) + (nSubsC * 9) + 1):((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9))
# indxEasy <- ((nSubsW + nSubsC + nSubsP) + 1):totalSubs #(nSubsW + nSubsC + nSubsP + nSubsE)
# indxEasyLong <- ((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9) + 1):((nSubsW * 9) + (nSubsC * 9) + (nSubsP * 9) + (nSubsE * 9))
# # the other, just summarizing for all participants
# # Not sure if this is necessary anymore
# # allProp <- data.frame(Group = c(rep("Wait",nSubsW*3),
# #                                 rep("Cognitive",nSubsC*3),
# #                                 rep("Physical",nSubsP*3)),
# #                       Handling = c(rep(2,nSubsW), rep(10,nSubsW), rep(14,nSubsW),
# #                                    rep(2,nSubsC), rep(10,nSubsC), rep(14,nSubsC), 
# #                                    rep(2,nSubsP), rep(10,nSubsP), rep(14,nSubsP)),
# #                       FivePts = rep(NA,(nSubsW + nSubsC + nSubsP)*9),
# #                       TenPts = rep(NA,(nSubsW + nSubsC + nSubsP)*9),
# #                       TwntyfivePts = rep(NA,(nSubsW + nSubsC + nSubsP)*9))
# # Get the proportion completed per combination for GLMs
# tempW <- propComplete()
# tempC <- propComplete(Data = cogData, nSubs = nSubsC, indx = indxCog, type = "Cognitive")
# tempP <- propComplete(Data = physData, nSubs = nSubsP, indx = indxPhys, type = "Physical")
# tempE <- propComplete(Data = easyData, nSubs = nSubsE, indx = indxEasy, type = "Easy")
# allProp2 <- rbind(tempW,
#                   tempC,
#                   tempP,
#                   tempE)  
# allProp2$totalComplete <- allProp2$propComplete * allProp2$totalTrials
# rm(tempW, tempC, tempP, tempE)
# # add optimal vector and difference from optimal
# wait <- c(rep(1,nSubsW*3),
#           c(rep(0,nSubsW), rep(1,nSubsW*2)),
#           c(rep(0,nSubsW*2), rep(1,nSubsW)))
# cog <- c(rep(1,nSubsC*3),
#           c(rep(0,nSubsC), rep(1,nSubsC*2)),
#           c(rep(0,nSubsC*2), rep(1,nSubsC)))
# phys <- c(rep(1,nSubsP*3),
#           c(rep(0,nSubsP), rep(1,nSubsP*2)),
#           c(rep(0,nSubsP*2), rep(1,nSubsP)))
# easy <- c(rep(1,nSubsE*3),
#           c(rep(0,nSubsE), rep(1,nSubsE*2)),
#           c(rep(0,nSubsE*2), rep(1,nSubsE)))
# allProp2$Optimal <- c(wait,cog,phys,easy)
# allProp2$Deviation <- allProp2$Optimal - allProp2$propComplete
# # sort by subject ID
# #allProp2 <- allProp2[order(allProp2$SubjID),]
# # actual proportions and standard errors, and the optimal summary for plotting
# trueProp <- aggregate(allProp2$propComplete,
#             by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
#             FUN = 'mean')
# trueSD <- aggregate(allProp2$propComplete,
#           by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
#           FUN = 'sd')
# colnames(trueProp) <- c("Reward","Handling","Group","propComplete")
# trueProp$SE[1:9] <- (trueSD$x[1:9])/(sqrt(nSubsW))
# trueProp$SE[10:18] <- (trueSD$x[10:18])/(sqrt(nSubsC))
# trueProp$SE[19:27] <- (trueSD$x[19:27])/(sqrt(nSubsP))
# trueProp$SE[28:36] <- (trueSD$x[28:36])/(sqrt(nSubsE))
# trueProp$Optimal <- aggregate(allProp2$Optimal,
#             by = list(allProp2$Reward, allProp2$Handling, allProp2$Group),
#             FUN = 'mean')$x

@@

# Plot coeffs
temp <- melt(logisRH[c("Handling", "Reward"), ])
qplot(data = temp, x = Var1, y = value, geom = "boxplot") + 
  labs(title = "Handling and Reward Coefficients", x = "", y = "Coefficients") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(legend.position = c(0.9, 0.7),
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.background = element_blank(), 
       axis.line = element_line(colour = "black"),
       text = element_text(size = 16))
```

@@

>>>>

---
title: "Preregistration analysis, Cost 2 study"
author: "Claudio Toro Serey & Joseph T. McGuire"
output:
  html_document: default
  pdf_document: default
---

@@

## 3. Research Questions

### 3.1.

Do decision makers in foraging environments integrate information about delay durations and reward amounts to produce reward-maximizing behavior?

### 3.2.

Main question: Are decisions affected differently by equivalent time periods of pure delay, cognitive effort, physical effort, and non-effortful task engagement? Do these four conditions involve different levels of subjective costs?

### 3.3.

How can we best computationally model the perceived cost of time and effort to predict choices in each condition?

## 4. Hypotheses

### 4.1.

Single-option, accept/reject decisions will be influenced by within-subject manipulations of reward magnitude and associated delay duration in line with a theoretical reward-maximizing strategy.

 **4.1.1.**	Participants will more frequently accept high-reward prospects than low-reward prospects.

@@

#### 4.2.

Prospect acceptance rates will differ across four between-subject conditions, in which the delays associated with rewards (a) are unfilled, (b) include a cognitive effort requirement, (c) include a physical effort requirement, or (d) require a trivial level of physical effort but have matched visual stimuli to the physical effort condition.

  **4.2.1.**	In the cognitive effort condition, overall acceptance rates will be greater than in the unfilled-delay condition.

@@

#### 4.3.

Choices will be well fit by a computational model in which the subjective opportunity cost of time is free to vary across the four between-subject conditions.

  **4.3.1.**	Participants will display stable preferences, meaning that the reward amounts they accept in a given timing condition will be similar throughout the experiment.

@@

## 5. Existing data

Registration prior to creation of data

## 6. Data collection

The study will use human subjects recruited from the university community, who will be recruited through the university job board. We will pay up to $30 for a session (1 hour), which includes incentive-based rewards. Participants will be ineligible if they have participated in our pilot study, or they are under 18 years of age. The study will run for however long it takes to reach our desired sample size.

## 7. Sample size

We will acquire a sample of 48 participants.

## 8. Sample size rationale?

Sample sizes were determined by means of power analysis (repeated measures ANOVA), using a significance level of 0.05, power of 0.8, a desired effect size of f = 0.5, and four groups (one for each condition). The resulting sample size was 45, which we increased to 48 in order to match the potential order of blocks and colors (section 15) while further ensuring that we have enough power.

## 9. Variables

### 9.1 The present study has three levels of manipulation.

9.1.1. Cost condition, manipulated within subjects. The conditions are (a) unfilled delay, (b) cognitive effort, and (c) physical effort. Unfilled delays will be paired with each effort type in a blocked manner (7-minute blocks). Each combination will be experienced three times.

9.1.2. Handling and travel times. These will dictate how long each trial is. The "handling time" governs how long it is necessary to wait or work to obtain a reward if a trial is accepted. The "travel time" governs how long the inter-trial interval will be, regardless of whether the trial is accepted. Participants will know that the handling time will be always 10 seconds, and the travel time 6 seconds (matching one of the between-group experiment conditions).

9.1.3. Reward magnitude, manipulated within-subject across individual trials. On each trial, participants will learn the amount of money they can earn if they complete the trial (i.e. 4, 8, or 20 cents). The amounts will be sampled uniformly, while guaranteeing that each participant experiences each amount at least twice during each block.

## 10. Measured variables

### 10.1. Behavioral variables

10.1.1. Choice (complete or quit a given trial).

10.1.2. Response times.

10.1.3. Total earnings.

10.1.3. Performance in the cognitive effort tasks (right or wrong answer).

10.1.4. Grip force throughout the experiment and maximum grip strength (physical effort condition).

### 10.2. Questionnaire

10.2.1. Age.

10.2.2. Gender.

10.2.3. Post-experimental oral assessment of choice strategy adoption. We will ask participants the following: "Did you follow a specific acceptance strategy? If so, what was it?" and "Which condition felt most effortful to you?". Answers will be written down by the experimenter, and recorded on an excel spreadsheet.

## 11. Indices

### 11.1. Behavioral data

11.1.1. Proportion of prospects accepted.

11.1.2. Percentage of cognitive effort trials with above-threshold errors (over 2 errors), in which participants are forced to travel to the next trial.

### 11.2. Optimal behavior

The theoretical opportunity cost of time will be calculated for each handling/travel/reward acceptance combination in order to assess choice optimality. This measure is computed as the cents per second possible when consistently accepting only certain reward amounts (e.g. 8 and 20 cents) in a given timing condition. This ground truth measure is independent of cost type, allowing us to assess subject optimality.

## 12. Study type

Experiment - A researcher randomly assigns treatments to study subjects, this includes field or lab experiments. This is also known as an intervention experiment and includes randomized controlled trials.

## 13. Blinding

No blinding is involved in this study.

## 14. Study design

The present study complements our previous one, which evaluated cost types through a between-group design. Many of the details remain the same. However, this time every participant will experience all forms of cost (delay, physical effort, and cognitive effort). Delay trials will be intermixed with both physical and cognitive trials, providing a common condition to compare these effortful costs. On each trial, participants will have the opportunity to either wait or work for 10 seconds (the "handling time") in order to earn a reward; alternatively, they can quit by pressing the space bar and skip to the next trial (which incurs a 6 second travel time). Two types of trial combinations (physical/wait and cognitive/wait) will be experienced three times by each participant in 7-minute-long, interleaved blocks. Reward amounts can be 4, 8, or 20 cents, uniformly distributed. Timing information will be disclosed at the beginning of each block, and reward prospects displayed before each trial begins.

## 15. Randomization

### 15.1. Block type 
Each block will contain unfilled delay trials and trials of one effortful cost type (se 14.1). Blocks with each cost type will be interleaved, and participants will be assigned to one of two orders: beginning with a cognitive/wait or a physical/wait block. A row will be assigned to subject IDs prior to recruitment, ensuring that each order is experienced equally often. Participants will be offered a break after completing three out of the six blocks.

### 15.2. Rewards and cost type 
Participants can earn 4, 8, or 20 cents per trial at the cost of performing one of the three conditions. These values are sampled without replacement from a vector containing every combination of cost (effort defined by the current block, and delay) and reward. The vector is reshuffled once completed. This ensures that each combination of cost and reward is presented at least twice per block.

### 15.3. Trial color 
During the offer phase, each cost type will appear in a unique color in order to ease its identification. Colors were chosen to contrast each other while being amenable to red-green color-blindness. Colors will be assigned to each cost type so that all three colors are equally paired to each cost type across participants. A row will be assigned to subject IDs prior to recruitment to indicate their block order and color pairing.

## 16. Analyses

@@

@@

# Plot coeffs
temp <- melt(logisRH[c("Handling", "Reward"), ])
qplot(data = temp, x = Var1, y = value, geom = "boxplot") + 
  labs(y = "Coefficients") +
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme(legend.position = c(0.9, 0.7),
       panel.grid.major = element_blank(), 
       panel.grid.minor = element_blank(), 
       panel.background = element_blank(), 
       axis.line = element_line(colour = "black"),
       text = element_text(size = 16))
```

@@

## 17. Inference criteria

These criteria will remain the same as before. For the ANOVAs, we will report the F statistic as well as the p-value. For the multiple linear regressions and logistic regression, we will report the betas (coefficients) and their associated p-values. When comparing two groups of values as in pairwise comparisons, we will use non-parametric permutation tests. We will report p-values and effect sizes, and all tests will be two-tailed. The alpha threshold for p-value evaluation will be set at 0.05.

## 18. Data exclusion

### 18.1. Missing data and quitting early.

Participants are told that they are allowed to stop participation in the study at any time without consequence. Data from participants who do not finish the experiment will be excluded from the study, and those participants will be replaced.

### 18.2. Inattentiveness

To guard against inattentiveness, a single 30 second catch trial will be placed at the end of each experimental block, merely prompting participants to press a key. From the pilot study, participants are able to respond within three seconds. Based on this, a response within three seconds will be considered as proper engagement, and a participant that fails two or more of these checks will be excluded and replaced.

### 18.3. Reasonable choices

The task is structured so that one reward amount must always be accepted. Thus, a participant who quits every trial in at least one block will be assumed not to have followed or understood task instructions, or to have disengaged from the task altogether, and will be excluded and replaced.

### 18.4. Performance

In the cognitive effort condition, participants were forced to travel if they made 2 mistakes in a trial. Any participant with more than 30% forced travels will be excluded and replaced.
